---
title: "Cleaning"
output: html_document
date: '2022-07-24'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T, fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(ggplot2, dplyr, tidyverse, gridExtra, ggrepel, plotly, skimr, tidytext, data.table) 
# install a package if it does not exist already and put the package in the path (library)
# dplyr, ggplot2,tidyr
```

# Dataset 
 
The online database Fbref includes a large database of soccer states in the English Premier League. Our group compiled every year's stats of the league table, standard stats table, goalkeeping stats, and shooting stats from the 2000-2001 season to 2021-2022 season

# Cleaning

```{r read in files}

for(year in 2000:2021) { #for loop through every year
  suffix <- paste0(year,".",(year+1)) #name changes based on each year
  suffix2 <- paste0(year,"-",(year+1))
  filepath <- paste0("data/",suffix2," - Sheet1.csv") 
  varname <- paste0("data", suffix)
  assign(varname, read.csv(filepath)) #assign read csv file to variable name
}
```

```{r add new columns}
#add year column and team year 
for(year in 2000:2021) { #for loop through every year
  suffix <- paste0(year,".",(year+1)) #name changes based on each year
  varname <- paste0("data", suffix)
  reference <- get(varname) #reference to actual variable with the same string name
  assign(varname, data.frame(Year = rep(year, 20), reference)) #assign read csv file to variable name
  reference <- get(varname)
  assign(varname, reference %>% mutate(reference, "Squad.Year" = paste(reference$Squad, reference$Year)))
}

```

```{r combine yearly datasets}
bigdata <- data2000.2001
for(year in 2001:2021) { #for loop through every year
  suffix <- paste0(year,".",(year+1)) #name changes based on each year
  varname <- paste0("data", suffix)
  reference <- get(varname) #reference to actual variable with the same string name
  bigdata <- merge(bigdata, reference, all = TRUE)
}
str(bigdata)
```

It is necessary to remove columns which are too directly correlated to points (Goals, Wins, Losses, Assists, etc), ones that are repeated twice, and also those that aren't useful at all. We are keeping those that are specified in the vector below.
```{r clean columns}
wantedcols <-  c("Pts","Attendance","X..Pl","Squad","Age","Poss","PK","PKatt","CrdY","CrdR","GA","SoTA","Saves","Save.","CS","CS.","PKatt","PK","PKsv","PKm","Sh","SoT","SoT.","Sh.90","SoT.90","G.Sh","G.SoT","Dist","PK.1","PKatt.1","Year","Squad.Year")

usefuldata <- bigdata[ , names(bigdata) %in% wantedcols] #too directly correlated info (goals and assists obviously)

usefuldata <- rename(usefuldata, Players.Used = X..Pl, Save_Pct = Save., CS_Pct = CS., SoT_Pct = SoT., G_Per_Sht = G.Sh, G_Per_SoT = G.SoT,PK_A = PK.1, PKatt_A = PKatt.1, Sh_Per_90= Sh.90, SoT_Per_90 = SoT.90)

 #better names
```



```{r sort table}
usefuldata <- usefuldata[order(usefuldata$Squad),] #order by squad alphabetically
str(usefuldata)
str( c("Squad","Year","Pts","Attendance","Players.Used","Age","Poss","PK","PKatt","CrdY","CrdR","GA","SoTA","Saves","Save_Pct","CS","CS_Pct","Pkatt","PK","PKsv","PKm","Sh","SoT","SoT_Pct","Sh_Per_90","SoT_Per_90","G_Per_Sht","G_Per_SoT","Dist","FK","PK_A","PKatt_A","Squad.Year"))

```

```{r graphs}
sct_plot1 <- usefuldata %>%
  ggplot(aes(x = Age, y = Pts)) + 
  # geometric options: color, size, shape, alpha: transparency (range: 0 to 1)
  geom_point(color = "blue", size= 3, alpha = .8) + 
  geom_text_repel(aes(label = Squad.Year), size = 3) +
  labs(title = "Premier League Team's Age vs Pts", 
       x = "Age", 
       y = "Point")
sct_plot1

sct_plot2 <- usefuldata %>%
  ggplot(aes(x = CrdY, y = Pts)) + 
  # geometric options: color, size, shape, alpha: transparency (range: 0 to 1)
  geom_point(color = "blue", size= 3, alpha = .8) + 
  geom_text_repel(aes(label = Squad.Year), size = 3) +
  labs(title = "Premier League Team's Yellow Cards vs Pts", 
       x = "Yellow Cards", 
       y = "Point")
sct_plot2

sct_plot3 <- usefuldata %>%
  ggplot(aes(x = CrdR, y = Pts)) + 
  # geometric options: color, size, shape, alpha: transparency (range: 0 to 1)
  geom_point(color = "blue", size= 3, alpha = .8) + 
  geom_text_repel(aes(label = Squad.Year), size = 3) +
  labs(title = "Premier League Team's Red Cards vs Pts", 
       x = "Red Cards", 
       y = "Point")
sct_plot3

sct_plot4 <- usefuldata %>%
  ggplot(aes(x = Save_Pct, y = Pts)) + 
  # geometric options: color, size, shape, alpha: transparency (range: 0 to 1)
  geom_point(color = "blue", size= 3, alpha = .8) + 
  geom_text_repel(aes(label = Squad.Year), size = 3) +
  labs(title = "Premier League Team's Save Percentage vs Pts", 
       x = "Save Percentage", 
       y = "Point")
sct_plot4

sct_plot5 <- usefuldata %>%
  ggplot(aes(x = Players.Used, y = Pts)) + 
  # geometric options: color, size, shape, alpha: transparency (range: 0 to 1)
  geom_point(color = "blue", size= 3, alpha = .8) + 
  geom_text_repel(aes(label = Squad.Year), size = 3) +
  geom_smooth(method="lm",formula = 'y~x', se = F, color = "red") +
  labs(title = "Premier League Team's Players Used vs Pts", 
       x = "Players Used", 
       y = "Point")
sct_plot5

pts_goals <- lm(Pts ~ GA, data = usefuldata)

ggplot(usefuldata, aes(x=GA,y=Pts)) + 
  geom_point() +
  geom_smooth(method="lm",formula = 'y~x', se = F, color = "red")

usefuldata <- usefuldata %>% select(-GA)

summary(pts_goals)

pts_goals <- lm(Pts ~ Attendance, data = usefuldata)
ggplot(usefuldata, aes(x=Attendance,y=Pts)) + 
  geom_point() +
  geom_smooth(method="lm",formula = 'y~x', se = F, color = "red")

usefuldata$Attendance <- as.numeric(gsub(",", "", usefuldata$Attendance))
summary(usefuldata$Attendance)
usefuldata$Poss <- as.numeric(gsub(",", "", usefuldata$Poss))
summary(usefuldata$Poss)
usefuldata$CrdY <- as.numeric(gsub(",", "", usefuldata$CrdY))
summary(usefuldata$CrdY)
usefuldata$CrdR <- as.numeric(gsub(",", "", usefuldata$CrdR))
summary(usefuldata$CrdR)
usefuldata$PKatt_A <- as.numeric(gsub(",", "", usefuldata$PKatt_A))
summary(usefuldata$PKatt_A)
usefuldata$Sh <- as.numeric(gsub(",", "", usefuldata$Sh))
summary(usefuldata$Sh)
usefuldata$SoT_Pct <- as.numeric(gsub(",", "", usefuldata$SoT_Pct))
summary(usefuldata$SoT_Pct)
usefuldata$Sh_Per_90 <- as.numeric(gsub(",", "", usefuldata$Sh_Per_90))
summary(usefuldata$Sh_Per_90)
usefuldata$G_Per_Sht <- as.numeric(gsub(",", "", usefuldata$G_Per_Sht))
summary(usefuldata$G_Per_Sht)
usefuldata$Dist <- as.numeric(gsub(",", "", usefuldata$Dist))
summary(usefuldata$Dist)
usefuldata$PKsv <- as.numeric(gsub(",", "", usefuldata$PKsv))
summary(usefuldata$PKsv)
usefuldata$PKm <- as.numeric(gsub(",", "", usefuldata$PKm))
summary(usefuldata$PKm)

write.csv(usefuldata, "data/usefuldata.csv", row.names = F)
```

```{r LASSO}
usefuldata <- read.csv("data/usefuldata.csv", header=T, na.string=c("", "?"))
dim(usefuldata)
names(usefuldata)
fit.data.lm <- lm(Pts~. -Squad.Year -Squad, usefuldata) # dump everything in the model
summary(fit.data.lm) 
```


``` {r adfs}
```
