---
jupyter:
  colab:
    name: EPL Neural Network (Failed)
  kernelspec:
    display_name: R
    name: ir
  language_info:
    name: R
  nbformat: 4
  nbformat_minor: 0
---

IMPORTANT
**NOTE: Unfortunately this model did not end up working, as the neural network did not train**
This .rmd **outlines the process** because it is important to understand even though it didnt work in the end.


# Machine Learning EPL (Keras)

First include all the necessary packages

``` {r echo = T}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(keras, ggplot2, glmnet, RColorBrewer, wordcloud, neuralnet, latex2exp, data.table, dplyr, tidyverse, gridExtra, ggrepel, plotly, skimr, tidytext,tensorflow) 

```


# Data Prepartion and Source

``` {r read, echo = T}
data <- read.csv("matches.csv")
#str(data)
```

Take out draws for now, because binary classification is enough for this simple NN example

``` {r binarization, echo = T}
data<-data[!(data$winner=="DRAW"),] #remove draws for now, converting this to binary classification problem
```

# Extra Data Cleaning

Make sure columns left are useful and also make sure the columns are results but predicting variables

```{r clean columns,echo = T, include=F}
removedcols <- c("date","match_name","winner","h_odd","a_odd","d_odd","team_winner","home_score","away_score","h_match_points","a_match_points")
data <- data[ , !names(data) %in% removedcols]

data <- data %>% 
  mutate(ls_winner = if_else(ls_winner == "HOME_TEAM",1,if_else(ls_winner=="AWAY_TEAM",0,2))) #string to binary format
```

#Read In Second Dataset

This dataset contains teamratings of attack, mid, and defense that will also be useful in determining match win
```{r combine other dataset echo = T}
big_df <- NULL
actyear <- 2005
years <- c("06","07","08","09","10","11","12","13","14","15","16","17","18","19","20","21") #shift one year to the right, because fifa years are wrong
for (yr in years) {
  filepath <- paste0("matchdata/fifa",yr," - Sheet1.csv")
  df <- read.csv(filepath)
  df$year <- actyear #season
  actyear <- actyear + 1
  big_df <- rbind(big_df, df) #bind together to one table
}

big_df <- big_df[!names(big_df) %in% c("League","Team.Rating")]
big_df <- big_df %>% rename(Squad = Name)
big_df <- big_df %>% rename(Year = year)
```

Unfortunately, the first dataset and this second dataset have slightly different names for the same teams, so we have to compare what they have at first and convert the changes later.

```{r find different columns}
other2 <- unlist(as.list(unique(big_df$Squad))) #these two different datasets have the same teams but also have names that are slightly different from each other i.e. Man Utd. vs Man United
other1 <- unlist(as.list(unique(data$home_team)))
sort(other1)
sort(other2)
```

Rename these in both dataframes so all the teams match

```{r renaming columns, echo = T}
#remaking teams the same team
data <- data.frame(lapply(data, function(x) {gsub("Utd", "United", x)}))

big_df <- data.frame(lapply(big_df, function(x) {gsub("Arsenal FC", "Arsenal", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Birmingham City", "Birmingham", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Blackburn Rovers", "Blackburn", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Bolton Wanderers", "Blackburn", x)})) 
big_df <- data.frame(lapply(big_df, function(x) {gsub("Brighton & Hove Albion" , "Brighton", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Cardiff City", "Cardiff", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Charlton Athletic", "Charlton", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Chelsea FC", "Chelsea", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Derby County", "Derby", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub(" FC", "", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Huddersfield Town", "Huddersfield", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Hull City", "Hull", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Leeds United", "Leeds", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Leicester City", "Leicester", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Newcastle United", "Newcastle", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Norwich City", "Norwich", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Queens Park Rangers", "QPR", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Stoke City", "Stoke", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Swansea City", "Swansea", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Tottenham Hotspur", "Tottenham", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("West Bromwich", "West Brom", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("West Bromwich Albion", "West Brom", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("West Ham United", "West Ham", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Wigan Athletic", "Wigan", x)}))
big_df <- data.frame(lapply(big_df, function(x) {gsub("Wolverhampton Wanderers", "Wolves", x)}))
```

# Combine Dataset

```{r combine datasets echo = T}
data <- full_join(data, big_df, by = c("home_team" = "Squad","season" = "Year"))  %>% 
  rename(ht_ATT = ATT, ht_MID = MID, ht_DEF = DEF, ht_OVR = OVR)

data <- full_join(data, big_df, by = c("away_team" = "Squad","season" = "Year"))  %>% 
  rename(at_ATT = ATT, at_MID = MID, at_DEF = DEF, at_OVR = OVR)

```

# Remove columns

Other columns could introduce bias

```{r remove unwanted columns echo = T}
teams <- c("home_team","away_team","season")
data <- data[ , !names(data) %in% teams]

more <- c("ht_rank")
data <- data[ , !names(data) %in% more]

data <- as.data.frame(sapply( data, as.numeric ))

data$ha_winner <- as.factor(data$ha_winner)
data$ls_winner <- as.factor(data$ls_winner)

str(data)
```
#Attempt at scaling data

```{r echo =T}
data$ht_ATT <- scale(data$ht_ATT)
data$ht_MID <- scale(data$ht_MID)
data$ht_DEF <- scale(data$ht_DEF)
data$ht_OVR <- scale(data$ht_OVR)
data$at_ATT <- scale(data$at_ATT)
data$at_MID <- scale(data$at_MID)
data$at_DEF <- scale(data$at_DEF)
data$at_OVR <- scale(data$at_OVR)
```
```{r echo=T}
#data <- mutate_all(data, function(x) as.numeric(as.character(x)))
```

```{r echo = T}
data <- data %>%
  select(ha_winner, everything())
#write.csv(data,"data.csv", row.names = FALSE)
```

### Generate training and testing datasets

**Reserve 4/5 entries for training data, and the rest for validation**

``` {r echo = T}
set.seed(10) #MESSI
n <- nrow(data)
portion1 <- as.integer(n*4/5)
testing.idx <- sample(n, portion1)

#Testing %>% select(-ha_winner)

ds <- data[testing.idx, -1] 
data_xtrain <- as.matrix(data[testing.idx, -1]) 
data_xtrain <- matrix(as.numeric(data_xtrain), ncol = ncol(data_xtrain))

data_ytrain <- as.matrix(data[testing.idx, 1] )
data_ytrain <- matrix(as.numeric(data_ytrain), ncol = ncol(data_ytrain))

#Training
data_xval <- as.matrix(data[-testing.idx, -1]) 
data_xval <- matrix(as.numeric(data_xval), ncol = ncol(data_xval))

data_yval <- as.matrix( (data[-testing.idx, 1] ))
data_yval <- matrix(as.numeric(data_yval), ncol = ncol(data_yval))

```

## Building the Neural Network

Experimenting with the number of layers


``` {r echo = T}
set_random_seed(7)
p <- dim(data_xtrain)[2]
model <- keras_model_sequential() %>%
  layer_dense(units = 45, activation = "relu", input_shape = c(p)) %>%  # 1 layer with 16 neurons. default activation is relu
  layer_dense(units = 1, activation = "softmax") # output

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy"), 
)
model
```

Compiling the fit

``` {r echo = T}
fit1 <- model %>% fit(
  (data_xtrain),
  (data_ytrain),
  epochs = 10,
  batch_size = 32,
  validation_split = .15
)
plot(fit1)
```

```{r echo = T}
weights <- model %>% get_weights()
str(weights) # show each layers of W's and b's
hist(weights[[2]])   # W^(1)
weights[[2]] # b's for layer 1
```

### **Assessing Performance**

What is the accuracy of this model?

What is its misclassification error?

``` {r echo = T}
results <- model %>% evaluate(data_xval, data_yval)
results
```

## Prediction

Predict the first 500 matches in the validation data and display labels

``` {r echo = T}
pred.prob <- model %>% predict(data_xval[1:500,])
pred.prob
y.pred <- model %>% predict(data_xval[1:500,])%>% k_argmax() %>% as.integer() #majority vote
vs.actual <- data.frame(yhat=y.pred, ydata=data_yval[1:500,1])
vs.actual <- vs.actual %>%
  mutate(correct = if_else(ydata == yhat,1,0))
vs.actual
```

**Unfortunately all yhat predicted values are 0 so clearly our Neural Network did not train!**