---
title: "NN With H2O R Library"
output: html_document
date: "2022-07-29"
---

# Machine Learning EPL (H2O)

The general purpose of this .rmd file is to use a neural network to predict match winner (exlcuding draws) in a soccer match in the English Premier League. All predictors are variables able to be calculated before a match starts such as win streak, average goals in past games, days since last match, etc. 

#Pre-resiquites

Import Library H2O Note: H2O requires a JRE (Java installed) from java.com before running. 
Also add other packages in case of need like dlpyr

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require('h2o')) {
  install.packages('h2o')
}
library(h2o)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(keras, ggplot2, glmnet, RColorBrewer, wordcloud, neuralnet, latex2exp, data.table, dplyr, tidyverse, gridExtra, ggrepel, plotly, skimr, tidytext,tensorflow) 
```

#Read Data 

data.csv is a cleaned data set ready for NN from EPL_ML.rmd (keras attempt at NN)

```{r echo =T}
m.data <- read.csv("NeuralNet_Data.csv")
h2o.init() #
```
Start h2o instance

```{r echo =T}
h2o.init() #
```

# Data Prepartation

Convert ha_winner to a factor (as it only has 0 and 1 as factor outputs)
Also split training and test data
```{r echo=T}
m.data$ha_winner <- as.factor(m.data$ha_winner)
tr <- m.data[-(3814:4814),]
test <- m.data[(3814:4814),]
```

# Train Neural Network

Create h2o instances of test and training data
```{r echo=T}
tr_h2o <- as.h2o(tr)
test_h2o <- as.h2o(test)
# Take a look at the contents of the H2OFrame
h2o.describe(tr_h2o)
```

Create Deep Learning NN Model
```{r echo=T}
h2o_nn <- h2o.deeplearning(x = 2:37, #36 cleaned predictors
                           y = 1,  #1 outcome
                           training_frame = tr_h2o,
                           nfolds = 5, 
                           standardize = TRUE,
                           activation = "Rectifier",
                           hidden = c(5,200), #2 hidden layers
                           seed = 10723, #seed doesnt make a difference
                           export_weights_and_biases = TRUE)
```

Store weights and biases of the 3 layers as data frame
```{r echo =T}
w1 <-as.data.frame( h2o.weights(h2o_nn,matrix_id=1))
w2 <- as.data.frame(h2o.weights(h2o_nn,2))
w3 <- as.data.frame(h2o.weights(h2o_nn,3))
b1 <- as.data.frame(h2o.biases(h2o_nn,1))
b2 <- as.data.frame(h2o.biases(h2o_nn,2))
b3 <-as.data.frame(h2o.biases(h2o_nn,3))

# Look at performance of model
h2o.performance(h2o_nn)
```

Predict match results based on testing dataset into dataframe
```{r echo =T}
# Use model to predict classes
h2o_predictions <- h2o.predict(object = h2o_nn,  newdata = test_h2o)

# Explore predictions
h2o_predictions <- as.data.frame(h2o_predictions)

```

```{r echo =T}
# Shutdown H2O instance
h2o.shutdown()
```

Calculate percent of match results that were predicted correct by our Neural Network model

```{r echo =T}
actual <- data.frame(yhat=h2o_predictions$predict, ydata=m.data[3814:4814,1])
actual <- actual %>% mutate(correct = if_else(ydata == yhat,1,0))#createa column for whether itis correct
actual <- actual[1:981,] #remove na values
v <- sum(actual$correct)
percent_pred <- sum(actual$correct)/(981) #percent correct
percent_pred

View(actual)
```

Calculate percent of match results that were predicted correct by OddsPortal Betting Odds

```{r echo=T}
sum <-sum( as.numeric(as.character(m.data$ha_winner[1:4794])))

act.data <- read.csv("matches.csv")
act.data<-act.data[!(act.data$winner=="DRAW"),] #remove draws for now, converting this to binary classification problem

act.data <- act.data %>% 
  mutate(bet_winner = (if_else(h_odd >= a_odd,0,1))) #string to binary format
act.data <- act.data %>% 
  mutate(bet_correct = (if_else(ha_winner == bet_winner,1,0))) #string to binary format


act.data <- act.data %>%
  select(bet_correct, everything())
act.data <- act.data %>%
  select(ha_winner, everything())
act.data <- act.data %>%
  select(bet_winner, everything())
act.data <- act.data %>%
  select(a_odd, everything())
act.data <- act.data %>%
  select(h_odd, everything())

bet.data <- act.data[,c(1:8)]
View(bet.data)

bets_sum <- sum(act.data$bet_winner)
bets_perc <- bets_sum / nrow(act.data)
bets_perc
```